---
output: 
    github_document:
        toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# ukbAid: Aid in doing analyses on the UK Biobank RAP

<!-- badges: start -->

[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

The goal of ukbAid is to help our research group at Steno Diabetes
Center Aarhus (SDCA) that is working on the UK Biobank (UKB) on the
research analysis platform (RAP).

This package and the general workflow we are using will help us be as
reproducible as possible, but without sharing any data that is not
permitted. It's a way for us to start moving in the direction of better
research practices.

## Installation

You can install the development version of ukbAid inside RStudio in the
R Console like so:

```{r}
# Install pak if not installed already.
# install.packages("pak")
pak::pkg_install("steno-aarhus/ukbAid")
```
