---
title: "Administration"
---

::: callout-warning
Still in development.
:::

## Adding users to the RAP project

1.  In the UKB AMS go to "Edit application", then go to the "Add member"
    form and select the scroll down menu, using either AUH or AU.
2.  In DNAnexus, go to `Project -> Share project` (people icon to top
    right, add by username or email address).
3.  After adding the users, delete their email from the survey sheet,
    since we don't need it.

## Adding users to the GitHub repos

1.  To add someone to the GitHub UKB team, use `gh_add_user_to_team()`.
2.  To add a contributor to a project, use the function
    `gh_add_user_to_repo()`.

## Todo after proposal submission

1.  Review the created Pull Request, provide any feedback or request
    clarifications if necessary.
    -   If changes necessary, main applicant should update the file on
        their own.

## Todo after project acceptance

1.  See a list of approved projects with `admin_get_project_ids()` and
    `gh_get_repos()` for projects on GitHub.
2.  Run `admin_start_approved_project()`, using the user name and
    project abbreviation to create the project, add it to the Steno
    GitHub, set up permissions, and add the user to the repo.

```{r}
#| eval: false
# Example of how to run the functions
# Find what the correct id is for the project
ukbAid:::admin_get_project_ids()
# Confirm the username to use
ukbAid:::gh_get_users()
# Then create the project!
ukbAid:::admin_start_approved_project("USER", "PROJECT_ID")
```

## General admin

### Google Forms to maintain submission to PR flow

We use Google Forms on a shared `ukbiobank-sdca` Drive folder for
researchers to submit project proposals. When a researcher submits a
project proposal, we have an "Apps Script" that takes the submission and
creates a Pull Request on the `ukbAid` repository. There is some
maintenance required for this workflow.

1.  Create a GitHub [fine-grained personal access
    token](https://github.com/settings/personal-access-tokens) with the
    following permissions on the "steno-aarhus" organization for the
    `ukbAid` repository (note: only someone with admin permissions for
    the organization can do this). Make sure this token only lasts for
    90 days.
    -   Read-only access to: "Contents" and "Metadata" (which is default
        and required)
    -   Read and write access to: "Actions", "Pull requests", and
        "Workflows"
2.  Copy the generated token and go into Google Drive. Open the
    `project-proposal-registration` Google Sheet (not the Form) and go
    to "Extensions -\> Apps Script". In this new page, on the line that
    says `var gh_token`, replace the string that starts with
    `github_pat_` with the token you just generated and copied. You may
    need to click "Run" to trigger the authorisation flow, but usually
    you don't need to.

You only need to do this flow whenever the token expires.

### On GitHub or own computer

Need to set a `SURVEY_ID` variable for the Google Forms Sheet ID (the
unique string that comes after the `/spreadsheets/d/` in the URL). This
needs to be set on GitHub as well as on your own computer if you are
going to download the survey data. This variable is used to download
the project proposal registration data, and you only need it (locally)
if you want to download the data to check something. 

1.  On GitHub, add a secret in Settings called `SURVEY_ID` with the
    Google Sheet ID.
2.  On your computer, you can either add a new variable in the form of
    `SURVEY_ID=""` in a `.Renviron` file or always get asked for it
    whenever you run the functions that require the variable (which uses
    the `ukbAid:::get_survey_id()` function that requires the ID).

### Uploading to Zenodo

For protocols, the code to use would be:

```{r}
#| eval: false
path <- here::here("path/to/protocol.qmd")
pdf_path <- zen_create_protocol_pdf(path)
client <- zen_create_file_record(
  path = pdf_path,
  metadata = zen_create_protocol_metadata(path),
  token = zen_get_token(),
  # Test it in the sandbox first
  sandbox = TRUE
)
```

Always test it in the sandbox first, which you can check that it uploads
by going to the [Zenodo sandbox](https://sandbox.zenodo.org/me/uploads).
