---
title: "Working in the Research Analysis Platform (RAP)"
execute: 
  eval: false
---

Because of the unique setup of the UKB RAP, the easiest way to do your
work is through Git and GitHub on the [SDCA GitHub
Organization](https://github.com/steno-aarhus/). This is due to a few
reasons:

1.  *Every* time you start up and enter into the UKB RAP, the working
    computer environment is *completely* clean... No personal files, no
    R packages installed, nothing. So we need some way of saving our
    work so we can download it again when we work again in the RAP.

2.  Whenever you finish working for the day, you *must* terminate your
    RStudio session in the RAP, because we pay for the RStudio use every
    hour. So we can't just leave the session running all the time, we
    have to turn it off. This forces us into the situation described in
    the point above.

3.  There are several of us working on this project and we want to be
    able to easily collaborate and help each other out. We will also
    have frequent code reviews of analyses. Both the collaboration and
    reviews are best done through Git and GitHub.

4.  The UKB organizing committee will review all project proposals and
    do some basic admin tasks, so keeping things organized and
    centralized on GitHub will help us be effective in our tasks.

5.  Because all code will be stored on the SDCA GitHub, it will be
    easier to share code and how we do things through it. So you the PhD
    student, postdoc, or researcher can better make use of all the
    efforts everyone is doing.

Because of these reasons, we expect everyone working on the Steno UK
Biobank project to use Git and use Steno's GitHub Organization account
to store their project.

## Steps *everytime* you enter into the UKB RAP

1.  Whenever you open up the UKB RAP, you won't have your project files
    nor have any packages installed. So you'll need to do a few set up
    tasks *each time* you work in the RAP. I've written a function that
    you can source directly that will take you through the steps needed
    to have everything set up. Run these two lines of code in the R
    Console, and afterwards follow the instructions on the Console.

    ```{r}
    #| eval: false
    install.packages("pak")
    pak::pak("steno-aarhus/ukbAid", ask = FALSE)
    ```

    Once this finishes, restart the session before running the next
    line:

    ```{r}
    #| eval: false
    ukbAid::setup_ukb_rap()
    ```

    ::: callout-note
    Since UKB RAP deletes everything when you Terminate the session,
    you'll be backing up your project on GitHub. What this function
    above does is install the necessary packages, installs the ukbAid
    package, sets up your Git config, sets up your authentication
    (credentials) to connect to GitHub, and finally downloads your
    GitHub repository into the RAP environment.
    :::

2.  Once the project is downloaded from GitHub and created the project
    in the RAP, you can open it by clicking the `.Rproj` file inside the
    project folder.

3.  You *may* have to update your credentials after opening up your
    project, so run this code to paste in your GitHub PAT token.

    ```{r}
    gitcreds::gitcreds_set()
    ```

4.  Then, run these two lines of code in the Console while inside your
    RStudio Project:

    ```{r}
    pak::pak(ask = FALSE)
    targets::tar_make()
    ```

    More details about this step is also found in your own project's
    main `README.md` file.

5.  When you've finished selecting the variables you want in the
    `data-raw/project-variables.csv`, you will need to open up the
    `data-raw/create-data.R` script and follow the instructions there to
    create your dataset.

Remember to frequently commit and push your file changes to your GitHub.
Otherwise all your work on the RAP will be lost next time you login.

## Other notes

-   **Save your changes in the Git history and push to your GitHub
    BEFORE Terminating your UKB-RAP project**: If you terminate before
    saving and pushing to GitHub, you *will* lose your work! Read the
    [Important notes on using GitHub](important-notes.qmd) page for more
    detail.

-   **Remember**, do not save any data in the Git history unless you
    have discussed it with the organizing committee. You can save
    results for tables and figures, as long as they are aggregate or
    statistical results.
